---
title: "RWorksheet_Garrido-3b"
author: "Angelo A. Garrido"
date: "2025-11-24"
output: pdf_document
---

## 1. Create a data frame using the table below
```{r}
#a. Write the codes.
Respondents <- 1:20
Sex = c(2,2,1,2,2,2,2,2,2,2,1,2,2,2,2,2,2,2,1,2)
Fathers_Occupation = c(1,3,3,3,1,2,3,1,1,1,3,2,1,3,3,1,3,1,2,1)
Persons_at_Home = c(5,7,3,8,5,9,6,7,8,4,7,5,4,7,8,8,3,11,7,6)
Siblings_at_School = c(6,4,4,1,2,1,5,3,1,2,3,2,5,5,2,1,2,5,3,2)

Types_of_Houses = c(1,2,3,1,1,3,3,1,2,3,2,3,2,2,3,3,3,3,3,2)
                    
respondents_data <- data.frame(
    Respondents,
    Sex,
    Fathers_Occupation,
    Persons_at_Home,
    Siblings_at_School,
    Types_of_Houses
)

```

```{r}
#b. Describe the data. Get the structure or the summary of the data
respondents_data

str(respondents_data)

summary(respondents_data)
```

```{r}
#c.Is the mean number of siblings attending is 5?
mean(Siblings_at_School)
#2.95, not 5
```


```{r}
#d. Extract the 1st two rows and then all the columns using the subsetting functions.
respondents_data[1:2, ]
respondents_data
```

```{r}
#e. Extract 3rd and 5th row with 2nd and 4th column.
respondents_data[c(3,5), c(2,4)]
respondents_data
```

```{r}
#f. Select the variable types of houses then store the vector that results as types_houses.
types_houses <- respondents_data$Types_of_Houses
types_houses
```

```{r}
#g. Select only all Males respondent that their father occupation was farmer.
all_Males <- subset(respondents_data, Sex == 1 & Fathers_Occupation == 1)
all_Males
```

```{r}
#h. Select only all females respondent that have greater than or equal to 5 number of siblings attending school.
all_Females <- subset(respondents_data, Sex == 2 & Siblings_at_School >= 5)
all_Females
```

## 2. Write a R program to create an empty data frame. Using the following codes:

```{r}
df = data.frame(Ints=integer(),
Doubles=double(), Characters=character(),
Logicals=logical(),
Factors=factor(),
stringsAsFactors=FALSE)
print("Structure of the empty dataframe:")
print(str(df))
```

## 3. Create a .csv file of this. Save it as HouseholdData.csv

```{r}
Respondents <- 1:10
Sex <- c("Male", "Female", "Female", "Male", "Male", "Female", "Female", "Male", "Female", "Male")
Fathers_Occupation <- c(1,2,3,3,1,2,2,3,1,3)
Persons_at_Home <- c(5,7,3,8,6,4,4,2,11,6)
Siblings_at_School <- c(2,3,0,5,2,3,1,2,6,2)
Types_of_Houses <- c("Wood","Congrete","Congrete","Wood","Semi-congrete","Semi-congrete",
                     "Wood","Semi-congrete","Semi-congrete","Congrete")

HouseholdData <- data.frame(
  Respondents,
  Sex,
  Fathers_Occupation,
  Persons_at_Home,
  Siblings_at_School,
  Types_of_Houses
)

HouseholdData


write.csv(HouseholdData, file = "HouseholdData.csv", row.names = TRUE )


data_set1 <- read.csv("HouseholdData.csv", header = TRUE)
data_set1
```

```{r}
#b. Convert the Sex into factor using factor() function and change it into integer.[Legend:
#Male = 1 and Female = 2]. Write the R codes and its output.

```
HouseholdData$Sex <- factor(HouseholdData$Sex, levels = c("Male", "Female"),labels = c(1, 2))

```{r}
#c. Convert the Type of Houses into factor and change it into integer. [Legend: Wood
#= 1; Congrete = 2; Semi-Congrete = 3]. Write the R codes and its output.

HouseholdData$Types_of_Houses <- factor(HouseholdData$Types_of_Houses,levels = c("Wood", "Congrete", "Semi-congrete"),labels = c(1, 2, 3))

HouseholdData$Types_of_Houses
```

```{r}
#d. On father’s occupation, factor it as Farmer = 1; Driver = 2; and Others = 3. What
#is the R code and its output?


HouseholdData$Fathers_Occupation <- factor(HouseholdData$Fathers_Occupation,levels = c(1, 2, 3),labels = c("Farmer", "Driver", "Others"))

HouseholdData$Fathers_Occupation
```

```{r}
#e. Select only all females respondent that has a father whose occupation is driver. Write
#the codes and its output.

female_driver <- subset(HouseholdData, Sex == 2 & Fathers_Occupation == "Driver")
female_driver
```

```{r}
#f. Select the respondents that have greater than or equal to 5 number of siblings attending
#school. Write the codes and its output.

many_siblings <- subset(HouseholdData, Siblings_at_School >= 5)
many_siblings


```

```{r}
#e. Extract 3rd and 5th row with 2nd and 4th column.
respondents_data[c(3,5), c(2,4)]
respondents_data
```

```{r}
#f. Select the variable types of houses then store the vector that results as types_houses.
types_houses <- respondents_data$Types_of_Houses
types_houses
```

```{r}
#g. Select only all Males respondent that their father occupation was farmer.
all_Males <- subset(respondents_data, Sex == 1 & Fathers_Occupation == 1)
all_Males
```

```{r}
#h. Select only all females respondent that have greater than or equal to 5 number of siblings attending school.
all_Females <- subset(respondents_data, Sex == 2 & Siblings_at_School >= 5)
all_Females

```

## 2. Write a R program to create an empty data frame. Using the following codes:

```{r}
df = data.frame(Ints=integer(),
Doubles=double(), Characters=character(),
Logicals=logical(),
Factors=factor(),
stringsAsFactors=FALSE)
print("Structure of the empty dataframe:")
print(str(df))

```

## 3. Create a .csv file of this. Save it as HouseholdData.csv

```{r}
Respondents <- 1:10
Sex <- c("Male", "Female", "Female", "Male", "Male", "Female", "Female", "Male", "Female", "Male")
Fathers_Occupation <- c(1,2,3,3,1,2,2,3,1,3)
Persons_at_Home <- c(5,7,3,8,6,4,4,2,11,6)
Siblings_at_School <- c(2,3,0,5,2,3,1,2,6,2)
Types_of_Houses <- c("Wood","Congrete","Congrete","Wood","Semi-congrete","Semi-congrete",
                     "Wood","Semi-congrete","Semi-congrete","Congrete")

HouseholdData <- data.frame(
  Respondents,
  Sex,
  Fathers_Occupation,
  Persons_at_Home,
  Siblings_at_School,
  Types_of_Houses
)

HouseholdData


write.csv(HouseholdData, file = "HouseholdData.csv", row.names = TRUE )


data_set1 <- read.csv("HouseholdData.csv", header = TRUE)
data_set1

```

```{r}
#b. Convert the Sex into factor using factor() function and change it into integer.[Legend:
#Male = 1 and Female = 2]. Write the R codes and its output.

HouseholdData$Sex <- factor(HouseholdData$Sex, levels = c("Male", "Female"),labels = c(1, 2))
```


```{r}
#c. Convert the Type of Houses into factor and change it into integer. [Legend: Wood
#= 1; Congrete = 2; Semi-Congrete = 3]. Write the R codes and its output.

HouseholdData$Types_of_Houses <- factor(HouseholdData$Types_of_Houses,levels = c("Wood", "Congrete", "Semi-congrete"),labels = c(1, 2, 3))

HouseholdData$Types_of_Houses
```

```{r}
#d. On father’s occupation, factor it as Farmer = 1; Driver = 2; and Others = 3. What
#is the R code and its output?


HouseholdData$Fathers_Occupation <- factor(HouseholdData$Fathers_Occupation,levels = c(1, 2, 3),labels = c("Farmer", "Driver", "Others"))

HouseholdData$Fathers_Occupation

```

```{r}
#e. Select only all females respondent that has a father whose occupation is driver. Write
#the codes and its output.

female_driver <- subset(HouseholdData, Sex == 2 & Fathers_Occupation == "Driver")
female_driver
```

```{r}
#f. Select the respondents that have greater than or equal to 5 number of siblings attending
#school. Write the codes and its output.

many_siblings <- subset(HouseholdData, Siblings_at_School >= 5)
many_siblings


```

## 4. Interpret the graph.
#Based on the results of the graph, negative opinion dominated the use of Twitter between July 14-21, 2020, with high peaks being on July 15 and 21. Throughout the week, the use of tweets that were neutral had a consistent frequency with positive tweets always bearing the lowest frequency. The data shows a clear picture of the platform that the negative thoughts had overwhelmed the positive thoughts considerably during the time.
