---
title: "Worksheet#3a"
author: "ANGELO A. GARRIDO"
date: "2025-10-13"
output: pdf_document
---
## 1. There is a built-in vector LETTERS that contains the uppercase letters of the alphabet and letters that contains the lowercase letters of the alphabet.
```{r}
#a. You need to produce a vector that contains the first 11 letters.
first_El_letters <- LETTERS[1:11]
first_El_letters

#b. Produce a vector that contains the odd-numbered letters
odd_num_letters <- LETTERS[seq(1, 26, by = 2)]
odd_num_letters

#c. Produce a vector that contains the vowel
vowels <- LETTERS[c(1, 5, 9, 15, 21)]
vowels

#d. Produce a vector that contains the last 5 lowercase letters.
last5 <- letters[22:26]
last5

#e. Produce a vector that contains letters between 15 to 24 (lowercase).
letters_Between15_24 <- letters[15:24]
letters_Between15_24
```


## 2. Create a vector(not a dataframe) with the average temperatures in April for Tugue garao City, Manila, Iloilo City, Tacloban, Samal Island, and Davao City. the average temperatures in Celcius are 42, 39, 34, 34, 30, and 27 degrees

```{r}
#a. Character vector for the cities

city <- c("Tuguegarao City", "Manila", "Iloilo City", "Tacloban", "Samal Island", "Davao City")
city

#b. Numeric vector for the temperature

temp <- c(42, 39, 34, 34, 30, 27)
temp

#c.Data Frame City and Temp
city_temp <- data.frame(city,temp)
print(city_temp)

#d. Rename the columns using names() function
names(city_temp) <- c("City", "Temperature")
city_temp

#e. Print the structure by using str() function. Describe the output.
str(city_temp)

#f. Display the content of row 3 and row 4
city_temp[3, ]
city_temp[4, ]

#g. Display the city with the highest and lowest temperature

# Highest temperature
city_temp[which.max(city_temp$Temperature), ]

# Lowest temperature
city_temp[which.min(city_temp$Temperature), ]


```
## 2. Create a matrix of one to eight and eleven to fourteen with four columns and threerows.

```{r}
#a. What will be the R code for the #2 question and its result?
matrix_data <- matrix(c(1:8, 11:14), nrow = 3, ncol = 4)
matrix_data

#result: matrix has 3 rows and 4 columns.numbers 1 to 8 and 11 to 14 are filled column-wise by default.


#b. Multiply the matrix by two. What is its R code and its result

matrix_data * 2

#result: Each Element in the matrix was multiplied by 


#c. What is the content of row 2? What is its R code

matrix_data[2, ]

#result: Row 2 contains the values 2, 5, 8, and 13.


#d. Display column 3 and column 4 in row 1 and row 2
matrix_data[1:2, 3:4]

#Output: 7, 12
#        8, 13


#e.Display only columns 2 and 3 in row 3
matrix_data[3, 2:3]

#Output: 6,11


#f. Display only column 4
matrix_data[, 4]

#Output:[1] 12 13 14


#g. Name the rows and columns

matrix_2 <- matrix_data * 2
rownames(matrix_2) <- c("isa", "dalawa", "tatlo")
colnames(matrix_2) <- c("uno", "dos", "tres", "quatro")
matrix_2

#Output: uno dos tres quatro
#isa       2   8   14     24
#dalawa    4  10   16     26
#tatlo     6  12   22     28

#h. Reshaping the Matrix

dim(matrix_data) <- c(6, 2)
matrix_data

#Output:
#  1   7
#  2   8
#  3   11
#  4   12
#  5  `13
#  6   14
```
## Using Arrays
## • Array can have more than two dimensions by using the array() function and dim()to specify the dimensions

## 3. An array contains 1, 2, 3, 6, 7, 8, 9, 0, 3, 4, 5, 1

```{r}
#a. Create a three-dimensional array
array_data <- array(rep(c(1, 2, 3, 6, 7, 8, 9, 0, 3, 4, 5, 1), 2),
                    dim = c(2, 4, 3))
array_data
#output: 


#     [,1] [,2] [,3] [,4]
#[1,]    1    3    7    9
#[2,]    2    6    8    0

#, , 2

#     [,1] [,2] [,3] [,4]
#[1,]    3    5    1    3
#[2,]    4    1    2    6

#, , 3
#     [,1] [,2] [,3] [,4]
#[1,]    7    9    3    5
#[2,]    8    0    4    1


#b. How many dimensions does the array have?

length(dim(array_data))
#output: 3


#c. Name the rows, columns, and array layers
rownames(array_data) <- c("a", "b")
colnames(array_data) <- c("A", "B", "C", "D")
dimnames(array_data) <- list(c("a", "b"),
                             c("A", "B", "C", "D"),
                             c("1st-Dimensional Array",
                               "2nd-Dimensional Array",
                               "3rd-Dimensional Array"))
array_data

#output:

#, , 1st-Dimensional Array

#  A B C D
#a 1 3 7 9
#b 2 6 8 0

#, , 2nd-Dimensional Array

#  A B C D
#a 3 5 1 3
#b 4 1 2 6

#, , 3rd-Dimensional Array

#  A B C D
#a 7 9 3 5
#b 8 0 4 1
```
